Thesis: draft

### 2021-05-04 draft.tex
worked on ch.4 real-world networks

It should be noted that not all real-world networks have a heterogeneous degree distribution. And even if the degree distribution is heterogeneous or fat-tailed, it does not always need to follow a power-law. The degree distribution can eg. also be a power-law with an exponential cut-off or a log-normal distribution.\\

@@

The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.\\
The conclusion of this section is that all the network models used in this thesis serve a goal and have specific topologies that are of interest in this thesis. But, it must always be kept in mind that, as the name itself already says, these models are just models of the real world. Results that are obtained from them should not be thoughtlessly extrapolated to real-world situations.

@@

Ideally, we would test the opinion dynamics model that was used throughout this thesis on network data directly obtained from social network companies. However, most of these companies are not to keen on sharing any form of data. This makes it very hard to find accurate and complete social media network data. Hence, we chose to test the model on other types of social networks, either on-line or off-line. Some possibilities are any form of friendship networks, e-mail networks, collaboration networks, etc. \\
\newline
One of the decisive properties of the real-world networks that were eventually used, was there size. The main, and actually only, reason for this was the computation time. If the networks would be to large, it would take weeks, sometimes even months before there would be any kind of results.\\
\newline
The size of the network models that were used, was around a thousand nodes. So for the real-world networks, we also used networks with a size of magnitude of thousand. This obviously benefits the computation time; it is, however, unknown how the network size affects the results. It is therefore recommended to test the model on larger network models and on larger real-world networks in future researches on this topic.

@@

In Chapter \ref{chap3} we already discussed quite a bit of opinion dynamics models and related topics such as algorithmic personalization, activation mechanisms and stubbornness. The goal of this section is to put every thing together and summarize the opinion dynamics model of this thesis.\\
\newline
This thesis follows the opinion dynamics model that was designed by Perra and Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''}, see reference \cite{Perra2019}. \\
\newline
We start with a network with $N$ users. Initially, each user is given one of the two opinions A or B. Each time step a random fraction equal to $p= 0.1$ of these users becomes active, where $p$ is the activation probability. As discussed in Chapter \ref{chap3}, Section \ref{actMech}, this thesis makes use of a constant activation probability. Every time step, all active users update their opinion synchronously. This updating is done as follows. First, the active user sends his current opinion to all his neighbors (active or not). Then, he updates his own opinion according to the opinions on his time-line. This updating follows the probabilistic majority model, as explained in Chapter \ref{chap3}, Section \ref{probMaj}. If another model instead of the probabilistic majority model is used, eg. the majority model, this will be explicitly stated. If not, one can assume that the probabilistic majority model is implemented. It is important to realize that what an active user sees on his/her time-line is only a selection of all the posts that his/her friends have posted. Remember from Chapter \ref{chap3}, Section \ref{actMech} that each user has a hidden time-line $L_i(t)$. This hidden time-line contains the time-line of the user at the last time he/she was active ($t_{\text{last}}$) plus all the post that his/her friends posted from $t_{\text{last}}$ until the user becomes active again at time $t$. If a user becomes active, the filtering algorithm (REF, REC or PR), discussed in Chapter \ref{chap3}, Section \ref{filter}, selects $S$ items from this hidden time-line and shows them on the actual time-line of the user. We will usually set $S = 20$, which quantifies the maximum length of the actual time-line.\\

@@

This model is a simple one, since it only allows for two different opinions in the system. The designed filtering algorithms are simple and intuitive as well. Some randomness is introduced by using the probabilistic majority model instead of the classic majority model. This might make the model a bit more realistic. This simpleness combined with the touch of randomness to give the model a more realistic vibe are definitely some strong points of using this model compared to other, more complex models. A simple model is more tractable and allows for a better understanding of what is going on in the dynamics. However, this might be a downside as well. Real-world opinion dynamics are, probably, far from simple. People do not update their opinion according to a simple probabilistic opinion dynamics model. Beside that, the cases where people have to choose between only two opinions are rather limited. Filtering algorithms, used by real social media companies, are unfortunately not publicly shared. We can however assume that they are not as simple as the ones we used in this thesis. Trying to model these far more complex systems by our very simple model, might lead to the fact that a lot of the dynamics might be missed or not well captured by our model. This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results. \\

>>>>

It should be noted that not all real-world networks have a heterogeneous degree distribution. And even if the degree distribution is heterogeneous or fat-tailed, it does not always need to follow a power-law. The degree distribution can eg. also be a power-law with an exponential cut-off, a log-normal distribution, a stretched exponential, etc \cite{Newman2005}.\\

@@

The fact that all the above models have a homogeneous degree distribution may also be a drawback, since most real-world networks have a heterogeneous degree distribution.

@@

Ideally, we would test the opinion dynamics model that was used throughout this thesis on network data directly obtained from social network companies. However, most of these companies are not to keen on sharing any form of data. This makes it very hard to find accurate and complete social media network data. Hence, we chose to test the model on other types of social networks, either on-line or off-line. Some possibilities are any form of friendship networks, e-mail networks, collaboration networks, etc. \\
\newline
One of the decisive properties of the real-world networks that were eventually used, was their size. The main, and actually only, reason for this was the computation time. If the networks would be to large, it would take weeks, sometimes even months before there would be any kind of results.\\
\newline
The size of the network models that were used, was around a thousand nodes. For the real-world networks, we also used networks with a size of magnitude of thousand to ten-thousand. This benefits the computation time; it is, however, unknown how the network size affects the results. It is therefore recommended to test the model on larger network models and on larger real-world networks in future researches on this topic. \\
\newline
We used two real-world networks. The data was obtained from \textit{The index of Complex networks} (ICON) \cite{ICON}. The first real-world network is an on-line friendship network among Finnish users of the social music listening and sharing website Last.fm (data from $2006$). Only the largest connected component of this on-line social network is included \cite{ICON}. This data set was originally used in reference \cite{Toivonen2009}, which describes the data set. This on-line social network has the following properties: number of nodes: $N = 8003$; number of edges: $16824$; modularity: $Q = 0.79$; average clustering coefficient: $\left<cc\right> = 0.31$; average degree: $\left<k\right> = 4.2$. The degree distribution is heterogeneous and is best fitted by a log-normal
\begin{equation}
	f(k) = \frac{1}{k \sigma \sqrt{2\pi}} \exp(-\frac{(\ln(k) - \mu)^2}{2 \sigma^2})\ ,
\end{equation}
with $\mu = 1.0078272$ and $\sigma = 0.95173779$. The degree distribution of the Last.fm social network, together with the log-normal fit, is shown in Figure \ref{deg_distr_lastfm}.\\
\newline
The second real-world network that was used is the giant component of the on-line, social network of users of the Pretty-Good-Privacy (PGP) algorithm for secure information interchange (data from $2004$) \cite{ICON}. This network was originally used in reference \cite{Boguna2004}, which gives a detailed description of the network data. The main properties of this network are: number of nodes: $N = 10680$; number of edges: $24340$; modularity: $Q = 0.85$; average clustering coefficient: $\left<cc\right> = 0.27$; average degree: $\left<k\right> = 4.56$. The network has a heterogeneous degree distribution, which is best fitted by two power-laws
\begin{equation}
 f(k) = a k^{-\gamma} \ ,
\end{equation}
with $a = 0.39604951$, $\gamma = 1.53568449$ for $k < 20$ and $a = 5.59021251$, $\gamma = 2.4944057$ for $k \geqslant 20$. The degree distribution for the PGP trust network, together with the power-law fits, is shown in Figure \ref{deg_distr_pgp}.

\begin{figure}[H]
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{Last.fm}
    \includegraphics[width=8cm, height=7cm]{lastfm_degree_distribution_8x7.png}
    \label{deg_distr_lastfm}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\textwidth}
    \caption{PGP}
    \includegraphics[width=8cm, height=7cm]{PGP_degree_distribution_8x7.png}
    \label{deg_distr_pgp}
  \end{subfigure}
  \captionsetup{format=plain}
  \caption[Degree distribution $P(k)$ of the two real-world networks used in this thesis.]{Degree distribution $P(k)$ of the two real world networks used in this thesis. \textbf{(a)} Last.fm on-line social network with a log-normal fit ($\mu = 1.0078272$, $\sigma = 0.95173779$); \textbf{(b)} PGP on-line trust network with two power-law fits ($a = 0.39604951$, $\gamma = 1.53568449$ for $k < 20$ and $a = 5.59021251$, $\gamma = 2.4944057$ for $k \geqslant 20$).}
\label{ev_op_50_50}
\end{figure}

@@

In Chapter \ref{chap3} we discussed quite a bit of opinion dynamics models and related topics such as algorithmic personalization, activation mechanisms and stubbornness. The goal of this section is to put everything together and summarize the opinion dynamics model of this thesis.\\
\newline
This thesis follows the opinion dynamics model that was designed by Perra and Rocha in their paper \textit{``Modelling opinion dynamics in the age of algorithmic personalization''} \cite{Perra2019}. \\
\newline
We start with a network with $N = 10^3$ users. Initially, each user is given one of the two opinions A or B. Each time step a random fraction equal to $p= 0.1$ of these users becomes active, where $p$ is the activation probability. As discussed in Chapter \ref{chap3}, Section \ref{actMech}, this thesis makes use of a constant activation probability. Every time step, all active users update their opinion synchronously. This updating is done as follows. First, the active user sends his current opinion to all his neighbors (active or not). Then, he updates his own opinion according to the opinions on his time-line. This updating follows the probabilistic majority model, as explained in Chapter \ref{chap3}, Section \ref{probMaj}. If another model is used, eg. the majority model, this will be explicitly stated. If not, one can assume that the probabilistic majority model is implemented. It is important to realize that what an active user sees on his/her time-line is only a selection of all the posts that his/her friends have posted. Each user has a hidden time-line $L_i(t)$. This hidden time-line contains the time-line of the user at the last time he/she was active ($t_{\text{last}}$) plus all the post that his/her friends posted from $t_{\text{last}}$ until the user becomes active again at time $t$. If a user becomes active, the filtering algorithm (REF, REC or PR), discussed in Chapter \ref{chap3}, Section \ref{filter}, selects $S$ items from this hidden time-line and shows them on the actual time-line of the user. We will usually set $S = 20$, which quantifies the maximum length of the actual time-line.\\

@@

This model is a simple one, since it only allows for two different opinions in the system. The designed filtering algorithms are simple and intuitive as well. Some randomness is introduced by using the probabilistic majority model instead of the classic majority model. This might make the model more realistic. This simpleness combined with the touch of randomness to give the model a more realistic vibe are definitely some strong points of using this model compared to other, more complex models. A simple model is more tractable and allows for a better understanding of what is going on in the dynamics. However, this might be a downside as well. Real-world opinion dynamics are, probably, far from simple. People do not update their opinion according to a simple probabilistic opinion dynamics model. Beside that, the cases where people have to choose between only two opinions are rather limited. Filtering algorithms, used by real social media companies, are unfortunately not publicly shared. We can however assume that they are not as simple as the ones we used in this thesis. Trying to model these far more complex systems by our very simple model, might lead to the fact that a lot of the dynamics might be missed or not well captured by our model. This downside of trying to model real-world social behavior with simple, tractable models is really a call for the urgent need for more real-world data to which the results obtained from simple models can be compared and which can help make the models capable of generating more realistic results. \\
